Trends of SDR in Thessaloniki

### 2023-11-05 MDPI_submition.Rmd
Commit 2023-11-05 13:09

We have demonstrated that in the period
`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is positive trend in SDR of for all-sky conditions ($`r alltrend`\,\%/y$).
A previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.

The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
The trend is also dependent on season, with winter showing the strongest statistically significant trend, in contrast to spring and summer.
Our findings are in agreement with other studies for the region.

$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$

>>>>

We have demonstrated that in the period
`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is positive trend in SDR of for all-sky conditions ($`r alltrend`\,\%/y$).
A previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$),suggests that cloud optical thickness is decreasing during this period.
Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we can not verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.

The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
The trend is also dependent on season, with winter showing the strongest statistically significant trend of
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
for all- and clear-skies, respectively, in contrast to spring and summer.
The trends for autumn are also significant but smaller (
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
for all- and clear-skies, respectively).
Our findings are in agreement with other studies for the region.

### 2023-11-05 MDPI_submition.tex
Commit 2023-11-05 13:09

We have demonstrated that in the period 1993 - 2023, there is positive
trend in SDR of for all-sky conditions (\(0.38\,\%/y\)). A previous
study \citep{Bais2013} for the period 1993 - 2011 the trend was found to
be \(0.33\,\%/y\). The increase of this trend, indicates that the
phenomenon and probably the causes, are still occurring. Accordingly, we
found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
further supports that the brightening can be independent of the cloud
occurrences, although, cloud formation can also attenuate SDR. We have
calculated the effect of TSI variability to be a negligible factor in
SDR variability. Unfortunately, there is no available data for this
period for the aerosols, in order to directly estimate their effect on
SDR, which is the most probable main cause. The attenuation of SDR by
aerosols over Europe have been proposed as major factor by
\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
SDR reaching the ground, and changes on their occurrences can effect the
net energy balance. Because we have no adequate data to investigate the
long term changes of cloud formation in the region, we can not infer if
the SDR trend we observe under cloud-sky is biased by changes in clouds
and local climate.

The observed brightening on SDR over Thessaloniki is dependent on SZA
(larger SZAs lead to stronger brightening). The trend is also dependent
on season, with winter showing the strongest statistically significant
trend, in contrast to spring and summer. Our findings are in agreement
with other studies for the region.

\(0.69\,\%/y\) \(0.83\,\%/y\)

>>>>

We have demonstrated that in the period 1993 - 2023, there is positive
trend in SDR of for all-sky conditions (\(0.38\,\%/y\)). A previous
study \citep{Bais2013} for the period 1993 - 2011 the trend was found to
be \(0.33\,\%/y\). The increase of this trend, indicates that the
phenomenon and probably the causes, are still occurring. Accordingly, we
found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
further supports that the brightening can be independent of the cloud
occurrences, although, cloud formation can also attenuate SDR. We have
calculated the effect of TSI variability to be a negligible factor in
SDR variability. Unfortunately, there is no available data for this
period for the aerosols, in order to directly estimate their effect on
SDR, which is the most probable main cause. The attenuation of SDR by
aerosols over Europe have been proposed as major factor by
\citet{Wild2021}. The dimming effect on SDR under cloudy-sky conditions
(\(-0.28\,\%/y\)),suggests that cloud optical thickness is decreasing
during this period. Because we have no adequate data to investigate the
long term changes of cloud thickness in the region, we can not verify if
the negative SDR trend we observe under under cloudy-skies can be
attributed solely to changes in clouds.

The observed brightening on SDR over Thessaloniki is dependent on SZA
(larger SZAs lead to stronger brightening). The trend is also dependent
on season, with winter showing the strongest statistically significant
trend of \(0.69\) and \(0.83\,\%/y\) for all- and clear-skies,
respectively, in contrast to spring and summer. The trends for autumn
are also significant but smaller ( \(0.43\) and \(0.37\,\%/y\) for all-
and clear-skies, respectively). Our findings are in agreement with other
studies for the region.
